---
title: "Anxiety EMA Main Analyses"
author: "Jihyun Hur"
date: "2025-05-19"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}

rm(list=ls())
library(readxl)
library(performance)
library(dplyr)
library(ggplot2)
library(tidyr)
library(tidyverse)
library(ggpubr)
library(broom)
library(robustbase)
library(ppcor)
library(lattice)
library(mediation)
library(ggeffects)
library(jtools)
library(sensemakr)
library(lme4)
library(nlme)
library(plm)
library(cowplot)
library(lmtest)
library(pracma)
library(mlVAR)
library(qgraph)
library(psych)
library(DHARMa)
library(glmmTMB)
library(MASS)
library(robustlmm)
library(brms)
library(patchwork)
library(emmeans)
library(effectsize)


# Set wd
cur_dir <- '/Users/jihyunhur/Yale/2_Github/project_anxiety_news/2_analysis'
dat_dir <- '/Users/jihyunhur/Yale/2_Github/project_anxiety_news/1_data'

# Load data
data <- data.frame(read_csv(file.path(dat_dir, 'project_anxiety_EMA_curated_data.csv')))

## Create summary data 
summary_data <- data %>% group_by(ID) %>% summarise(baseline_anxiety_score = unique(baseline_anxiety_score),
                                                    mean_anxiety_score = unique(mean_anxiety_score),
                                                    mean_info_seek = unique(mean_info_seek),
                                                    mean_brood = mean(raw_daily_brood_sum_score, na.rm=T),
                                                    mean_dampen = mean(raw_daily_dampen_sum_score, na.rm=T))
## Factorize gender group
data$gender <- factor(data$gender,levels = c("1","2","3"), labels = c("Male", "Female", "Nonbinary"))
data$race_category <- factor(data$race_category)

```

## H1
### Correlation between anxiety and information-seeking
```{r h1}
cor.test(summary_data$baseline_anxiety_score, summary_data$mean_info_seek)

cor.test(summary_data$mean_anxiety_score, summary_data$mean_info_seek)
```

## H2
### Multilevel Regression: Daily Anxiety ~ Daily Information-Seeking
```{r h2}
ctrl <- lmeControl(opt='optim')

h2_base_model <- lme(daily_anxiety_score ~ 
                       time + person_centered_info_seek + daily_info_seek + 
                       race_category + current_treat,
                     random = ~ 1 + daily_info_seek | ID, 
                     control=ctrl, 
                     data = data, 
                     na.action = na.exclude)

# Table 2
summary(h2_base_model)
parameters::model_parameters(h2_base_model, standardize = "refit")

h2_full_model <- lme(daily_anxiety_score~ 
                       time + person_centered_info_seek + daily_info_seek + 
                       race_category + current_treat +
                       age + gender + edu,
                     random = ~ 1 + daily_info_seek | ID, 
                     control=ctrl, 
                     data = data, 
                     na.action = na.exclude)

# Supplementary Table 1
summary(h2_full_model)
parameters::model_parameters(h2_full_model, standardize = "refit")

```

## H3-1
### Multilevel Regression: Daily Anxiety ~ Daily Emotion Regulation*Daily Information-Seeking
#### Brooding
```{r h3-1 brooding MS}
ctrl <- lmeControl(opt='optim')

h3_brood_base_model <- lme(daily_anxiety_score ~ 
                       time + person_centered_info_seek + person_centered_brood_sum_score +
                       daily_info_seek*daily_brood_sum_score +
                       race_category + current_treat,
                     random = ~ 1 + daily_info_seek + daily_brood_sum_score | ID, 
                     control=ctrl, 
                     data = data,
                     na.action = na.exclude)

# Table 3 - Brooding
summary(h3_brood_base_model)
parameters::model_parameters(h3_brood_base_model, standardize = "refit")

# Figure 1
## Figure 1A
ms_fig_1a_predict <- ggpredict(h3_brood_base_model, c("daily_info_seek", "daily_brood_sum_score[meansd]"))

ms_fig_1a <- plot(ms_fig_1a_predict) + xlab('Daily COVID-19 Information-Seeking') + 
  ylab('Daily Anxiety Score') +
  scale_colour_brewer(palette = "Set1", labels = c("-1 SD", "Mean", "+1 SD")) + 
  scale_fill_brewer(palette = "Set1") +
  theme_classic() + theme(legend.position = "top") + 
  ggtitle("") + labs(colour = "Daily Brooding") +
  theme(plot.margin = margin(t = 10, r = 5, b = 5, l = 5))

## Figure 1B
ms_fig_1b_predict <- ggeffects::ggpredict(h3_brood_base_model, c("daily_info_seek"))
ms_fig_1b <- plot(ms_fig_1b_predict) + xlab('Daily COVID-19 Information-Seeking') + 
  ylab('Daily Anxiety Score') + 
  geom_point(aes(color = "dummy_legend_label"), alpha = 0, size = 0) +
  theme_classic() + 
  # Below code is mainly for formatting
  scale_color_discrete(
    name   = " ",                     
    breaks = "dummy_legend_label",   
    labels = " ",                      
    guide  = guide_legend(
      override.aes = list(
        shape = NA,   
        color = NA,    
        fill  = NA,    
        alpha = 0     
      )
    )
  ) + 
  ggtitle("") + theme(legend.position = "top")

ggarrange(ms_fig_1a, ms_fig_1b, ncol = 2, nrow=1)

# Simple Slope Analysis
zvals <- c(-1.53, 0, 1.53) 
emtrends(h3_brood_base_model, specs="daily_brood_sum_score", var ="daily_info_seek", at = list(daily_brood_sum_score = zvals)) |> test()

# Correlation: mean info-seek ~ mean brooding
cor.test(summary_data$mean_info_seek, summary_data$mean_brood)

```

```{r h3-1 brooding SI}
ctrl <- lmeControl(opt='optim')

h3_brood_full_model <- lme(daily_anxiety_score ~ 
                             time + person_centered_info_seek + person_centered_brood_sum_score +
                             daily_info_seek*daily_brood_sum_score +
                             race_category + current_treat +
                             age + gender + edu,
                     random = ~ 1 + daily_info_seek + daily_brood_sum_score | ID, 
                     control=ctrl, 
                     data = data, 
                     na.action = na.exclude)

# Table S2 - Brooding
summary(h3_brood_full_model)
parameters::model_parameters(h3_brood_full_model, standardize = "refit")

# Figure S2
data_fig <- data %>%
  mutate(daily_brood_sum_score_group = case_when(
    daily_brood_sum_score < -1.53 ~ "-1 SD",
    daily_brood_sum_score >= -1.53 & daily_brood_sum_score < 1.53 ~ "Mean",
    daily_brood_sum_score >= 1.53 ~ "+1 SD"
  )) %>% filter(!is.na(daily_brood_sum_score_group))

data_fig$daily_brood_sum_score_group <- factor(data_fig$daily_brood_sum_score_group, levels = c("-1 SD", "Mean", "+1 SD"))

data_fig %>%
  ggplot(aes(x = daily_info_seek, y = daily_brood_sum_score)) +
  geom_point(
    alpha = 0.5,
    aes(shape = daily_brood_sum_score_group, color = daily_brood_sum_score_group),
    size=2
  ) + 
  ylim(c(-11, 12)) +
  scale_colour_brewer(palette = "Set1", labels = c("-1 SD", "Mean", "+1 SD")) +
  scale_shape_manual(
    values = c("-1 SD" = 16, "Mean" = 17, "+1 SD" = 15),  
    labels = c("-1 SD", "Mean", "+1 SD"),
    na.translate = FALSE
  ) +   
  theme_pubr() +
  theme(legend.position = "right") +  
  labs(
    colour = "Daily Brooding",
    shape = "Daily Brooding"
  ) + 
  geom_hline(yintercept = 0, linewidth = 0.8, linetype = 2) +
  geom_vline(xintercept = 0, linewidth = 0.8, linetype = 2) +
  ylab("Daily Brooding (Centered)") +
  xlab("Daily Information-Seeking (Centered)")

# Correlation: daily info-seek ~ daily brooding
cor.test(data$daily_info_seek, data$daily_brood_sum_score)

```

## H3-2
### Multilevel Regression: Daily Anxiety ~ Daily Emotion Regulation*Daily Information-Seeking
#### Dampening
```{r h3-2 dampening MS}
ctrl <- lmeControl(opt='optim')

h3_dampen_base_model <- lme(daily_anxiety_score ~ 
                       time + person_centered_info_seek + person_centered_dampen_sum_score +
                       daily_info_seek*daily_dampen_sum_score +
                       race_category + current_treat,
                     random = ~ 1 + daily_info_seek + daily_dampen_sum_score | ID, 
                     control=ctrl, 
                     data = data, 
                     na.action = na.exclude)

# Table 3 - Dampening
summary(h3_dampen_base_model)
parameters::model_parameters(h3_dampen_base_model, standardize = "refit")

# Figure S3
## Figure S3A
ms_fig_s3a_predict <- ggeffects::ggpredict(h3_dampen_base_model, c("daily_info_seek", "daily_dampen_sum_score[meansd]"))

ms_fig_s3a <- plot(ms_fig_s3a_predict) + xlab('Daily COVID-19 Information-Seeking') + 
  ylab('Daily Anxiety Score') +
  scale_colour_brewer(palette = "Set1", labels = c("-1 SD", "Mean", "+1 SD")) + 
  scale_fill_brewer(palette = "Set1") +
  theme_classic() + theme(legend.position = "top") + 
  ggtitle("") + labs(colour = "Daily Dampening") +
  theme(plot.margin = margin(t = 10, r = 5, b = 5, l = 5))


## Figure S3B
ms_fig_s3b_predict <- ggeffects::ggpredict(h3_dampen_base_model, c("daily_info_seek"))
ms_fig_s3b <- plot(ms_fig_s3b_predict) + xlab('Daily COVID-19 Information-Seeking') + 
  ylab('Daily Anxiety Score') + 
  geom_point(aes(color = "dummy_legend_label"), alpha = 0, size = 0) +
  theme_classic() + 
  # Below code is mainly for formatting
  scale_color_discrete(
    name   = " ",                     
    breaks = "dummy_legend_label",   
    labels = " ",                      
    guide  = guide_legend(
      override.aes = list(
        shape = NA,   
        color = NA,    
        fill  = NA,    
        alpha = 0     
      )
    )
  ) + 
  ggtitle("") + theme(legend.position = "top")

ggarrange(ms_fig_s3a, ms_fig_s3b, ncol = 2, nrow=1)

# Simple Slope Analysis
zvals <- c(-1.45, 0, 1.45) 
emtrends(h3_dampen_base_model, specs="daily_dampen_sum_score", var ="daily_info_seek", at = list(daily_dampen_sum_score = zvals)) |> test()

```


```{r h3-2 dampening SI}
ctrl <- lmeControl(opt='optim')

h3_dampen_full_model <- lme(daily_anxiety_score ~ 
                             time + person_centered_info_seek + person_centered_dampen_sum_score +
                             daily_info_seek*daily_dampen_sum_score +
                             race_category + current_treat +
                             age + gender + edu,
                     random = ~ 1 + daily_info_seek + daily_dampen_sum_score | ID, 
                     control=ctrl, 
                     data = data, 
                     na.action = na.exclude)

# Table S2 - Dampening
summary(h3_dampen_full_model)
parameters::model_parameters(h3_dampen_full_model, standardize = "refit")

```

## eH4
### Time-Lagged Multilevel Regression: Daily Anxiety (day t) ~ Daily Emotion Regulation (day t-1)*Daily Information-Seeking (day t-1)
#### Brooding
```{r eh4 brooding SI}
ctrl <- lmeControl(opt='optim')

# Add lagged data
lagged_function <- function(dat, col_name, org_col) {
  subjs <- unique(dat$ID)
  dat <- dat %>%
    mutate(time_numeric = as.numeric(stringr::str_extract(time, "\\d+")))
  
  dat <- dat %>%
    group_by(ID) %>%
    arrange(ID, survey_number) %>%
    # Create a column to check if times are consecutive
    mutate(
      time_diff = c(NA, diff(survey_number, lag = 1)),
      # Only assign lagged value if the time difference is 1 (consecutive)
      !!col_name := ifelse(time_diff == 1, dplyr::lag(.data[[org_col]], 1), NA)
    ) %>%
    # Remove the helper columns
    dplyr::select(-time_diff, -time_numeric) %>%
    ungroup()
  
  return(dat)
}

cols <- c("daily_brood_sum_score", "daily_dampen_sum_score", "daily_info_seek", "daily_anxiety_score")

for (col in cols) {
  new_col_name <- paste0(col, "_lagged", sep="")
  data <- lagged_function(data, new_col_name, col)
}


eh4_brood_base_model <- lme(daily_anxiety_score ~ daily_anxiety_score_lagged +
                             time + person_centered_info_seek + person_centered_brood_sum_score +
                             daily_info_seek_lagged*daily_brood_sum_score_lagged +
                             race_category + current_treat,
                     random = ~ 1 + daily_info_seek_lagged + daily_brood_sum_score_lagged  | ID, 
                     control=ctrl, 
                     data = data, 
                     na.action = na.exclude)

# Table S3 - Brooding
summary(eh4_brood_base_model)
parameters::model_parameters(eh4_brood_base_model, standardize = "refit")

```
## Dampening
```{r eh4 dampening SI}
ctrl <- lmeControl(opt='optim')

eh4_dampen_base_model <- lme(daily_anxiety_score ~ daily_anxiety_score_lagged +
                             time + person_centered_info_seek + person_centered_dampen_sum_score +
                             daily_info_seek_lagged*daily_dampen_sum_score_lagged +
                             race_category + current_treat,
                     random = ~ 1 + daily_info_seek_lagged + daily_dampen_sum_score_lagged  | ID, 
                     control=ctrl, 
                     data = data, 
                     na.action = na.exclude)

# Table S3 - Brooding
summary(eh4_dampen_base_model)
parameters::model_parameters(eh4_dampen_base_model, standardize = "refit")

```
